---
title: "Lab 10 Nate"
author: "Nathaniel Burola"
date: "December 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab 10 Objectives:

- Correlation matrices
- Binary logistic regression
- Dates/times with lubridate
- Heatmap of solar irradiation in Santa Barbara, Hilo (HI), and Fairbanks (AK) in 2010
- Aquaculture with gganimate

# PART 0. Load libraries
```{r packages}

library(tidyverse) # The tidyverse!
library(corrplot) # For correlation matrices
library(janitor) # For cleaning up column names
library(lubridate) # For dealing with dates & times
library(gganimate) # For adding animation to ggplot graphics
library(ggimage) # For updating graph components with images


```

# PART 1. Correlation matrix (World Bank world environmental factors data)

Compiled World Bank data, accessed from: https://www.kaggle.com/zanderventer/environmental-variables-for-world-countries#World_countries_env_vars.csv

- Read in the "world_env_vars.csv file" as env_var
- Find correlations between all variables, columns 2:28
- Plot the correlation matrix (corrplot)

```{r env_corr}
env_var <- read.csv("world_env_vars.csv") %>% 
  na.omit #Omit the N/A values (NEVER DO THIS UNLESS YOU EXPLORE THE DATA FIRST)

cor_df <- cor(env_var[2:28]) #Plotting correlation values for coloumns 2-28 from -1 to 1 (remember if it is -1 or 1 it is plotting against itself)

corrplot(cor_df, 
         type = "upper", 
         method = "ellipse", 
         tl.col = "black", 
         tl.cex = 0.5)

#Big red values are strongly negatively correlated 
#Big blue values are strongly postively correlated 
#Amount of rain and the driest month perfectly correlated with the amount of rain and the driest quarter which is postively strongly correlataed
#Isothermality and temperature seasonality which is negatively strongly correlated 

```

# PART 2. Binary Logistic Regression (Donner Party Data)

Use the 'glm' function for fitting *generalized linear models* (the *logit* - log odds of survival, in our case, will be linearly related to Sex and Age. So we expect the final model to look something like this: 

$$Log Odds (Survival) = \beta_0 + \beta_1(Age) + \beta_2(Sex)$$

We'll use 'family = binomial' to run binomial logistic regression...otherwise, this looks very similar to other types of regression we've already done. 

a. Read in the DonnerTable.csv file as DonnerTable

```{r donner_data}

#Survival is the single outcome variable, not a continous measured variable, and is based on age and sex. 

donnertable <- read_csv("DonnerTable.csv")

#Explicity putting your outcome table as 0s and 1s (cocercing)
#Survival is 1 
#Dying is 0 
```

b. Binomial logistic regression
1 =  Survival 
0 = Death
What is the probability of that outcome based on a Donner party's member sex and age?

```{r blr}

donner_blr <- glm(Outcome ~ Sex + Age, family = "binomial", data = donnertable)

summary(donner_blr)

#Generalized Linear Model with something on the right (log series) that is linearly related to the left (linear series)
#Postive number is going up 
#Negative number is going down 
#This model for logisitic regression that the odds of female surviving is higher than the odds of male surviving 
#Odds of surviving decreases as age increases 
#Survival is better for kids and women that stayed at Donner Lake compared to the men that went out for help (realistically)

```

c. Use the model to answer: What are the log odds of survival for a 5 year-old female? The probability of survival?

In order to use the predict function, you need to create an exact same new data frame (sex and age written exactly the same)
for bionomial logisitic regresion 

```{r predict_f_5}

# 1. Create a data frame with variables Sex and Age, containing data "Female" and 5, respectively: 

f_5 <- data.frame(Sex = "Female", Age = 5) #Creating a new data frame with sex and 

# 2. Find the log odds of survival for the new data (5 year old female) using predict() function with type = "link":

f_5_logodds <- predict(donner_blr, newdata = f_5, type = "link")

f_5_logodds #Log odds of survival is 1.443775 


# 3. Exponentiate the log odds to find ODDS of survival for a 5 year old female:

exp(f_5_logodds) #Exponentiate the log odds for a 5 year old girl is 4.23666


# Ask: Could we manually find the probability of survival for a 5-year old female? recall: p/(1-p) = ODDS

#Answer is 81% chance of survival for a 5 year old female in the party 

f_5_prob <- predict(donner_blr, newdata = f_5, type = "response")

f_5_prob

# 4. Actually, let's just use type = "response" in the predict function, which converts to a probability for us:



```

d. What is the probability of survival for a 25 year-old male? 

```{r predict_m_25}

# Similarly:

m_25 <- data.frame(Sex = "Male", Age = 25) # Make a new data frame

m_25_prob <- predict(donner_blr, newdata = m_25, type = "response") # Find probability of survival
m_25_prob

#Probability of a 25 year old male surviing is 42% (half compared to the 5 year old of 84%)

```

e. Create new sequences of data so that we can graph probabilities for the entire spectrum of ages, designated by sex.

```{r new_data}

#WARNING: CODE WILL APPEAR ON THE FINAL 

seq_age <- rep(seq(from = 0, to = 100), 2) # Create a vector sequence from 0 to 100, twice (one will be "Male" and one will be "Female") will create a sequence of ages for both female and male 

f_101 <- rep("Female", 101) # Repeat 'Female' 101 times (to match years data) (0-100 in incriments of 1 have 101 values)
m_101 <- rep("Male", 101) # Repeat 'Male' 101 times
mf_101 <- c(f_101, m_101) # Combine them into a single vector

# Combine the age and sex sequences into a single data frame - that will be the new data that we have our model make predictions for

donner_newdata <- data.frame(seq_age, mf_101) # MUST make column names match variables in the model!
colnames(donner_newdata) <- c("Age","Sex") #Make these column names the exact same names as the ones in the original data set 

donner_newdata

```

f. Now that we have new data to put into our model to have it make predictions, let's go ahead and actually find the predicted probabilities for each Age/Sex combination.

```{r donner_predict}

# Find probabilities using predict (with type = "response"). Include SE.

predicted_probs <- predict(donner_blr, newdata = donner_newdata, type = "response", se.fit = TRUE)


# Coerce outcome into data frame. 

graph_data <- data.frame(donner_newdata, predicted_probs$fit, predicted_probs$se.fit)

colnames(graph_data) <- c("Age", "Sex", "Probability", "SE")

graph_data #This in Tidy format by the way...
```

g. Graph results.

```{r donner_graph}
#Will make a line graph and a ribbon graph that will extend up and down from the predictive line (top of the ribbon and bottom of the ribbon will be -1 standard deviation and 1 standard deviation)

#Males have a lower probability of survival than females 
#Probability of survival goes down as age increases 
#Creating  a geom_ribbon for minimum value for ribbon will be probability column minus 1 standard error (SE)
#Maximum value for ribbon will be probability column plus 1 standard error (SE)
#Finalizing graph will be better with y-axis from 0-1

ggplot(graph_data, aes(x = Age, y = Probability)) + 
  geom_line(aes(color = Sex)) + 
  geom_ribbon(aes(ymin = Probability - SE, ymax = Probability + SE, fill = Sex, alpha = 0.5))
  
```


# PART 3. Solar irradiation at the 3 locations in 2010 (pull in all together, do some lubridate stuff, etc.)

a. Read in multiple solar irradiation files (for SB (CA), Hilo (HI), and Fairbanks (AK)):
```{r solar_data}

si_full <- list.files(pattern = "solar_irradiation_*") %>% 
  map_df(~read_csv(.)) %>%  
  clean_names()
#Searching for any files names that begin with solar_irradiation
#Read all the files that have their names starting with solar_irradiation and add them together (SB, HI, AK)
#clean_names needs the janitor package and will line up variables as best it can 

si_full

```

b. Wrangle the data
```{r wrangle_solar}

solar_tidy <- si_full %>%  
  rename(
    sol_rad = etr_wh_m_2,
    date = yyyy_mm_dd,
    time = hh_mm_lst
    ) %>%  
  filter(time != "NA") %>%  
  mutate(site = fct_relevel(site, "Hawaii", "Santa Barbara", "Alaska"))

solar_tidy




```

c. Use lubridate() functions to convert to times/dates
```{r date_time}

solar_tidy$date <- mdy(solar_tidy$date) #Based on the values in the column, lubridate will infer what the values are, and will organize them in a format

solar_tidy$time <- hms(solar_tidy$time)

```

d. Make an awesome figure of solar irradiation (heat/tile map)
```{r solar_graph}

solar_gg <- ggplot(solar_tidy, aes (x = date, y = time)) + 
  geom_tile(aes(fill = sol_rad)) + 
  scale_fill_gradientn(colors = c("royalblue2", "mediumorchid1", "orange", "yellow")) +
  scale_y_time() +
  facet_grid(site ~ .)

solar_gg



```


#PART 4. gganimate example: total aquaculture production (metric tons) for United States, Brazil, Chile, and Ecuador

a. Get the data, and do some wrangling:
```{r aqua_data}

aq_df <- read_csv("aq_wb.csv")

aq_df

aq_tidy <- aq_df %>% 
  filter(country == "Brazil" |
           country == "Chile" |
           country == "Ecuador" |
           country == "United States") %>% 
  gather(year, aq_prod, `1960`:`2016`) %>% 
  filter(year >= 1990) %>%  
  mutate(aq_mil = aq_prod/1000000) %>% 
  select(country, year, aq_mil)

aq_tidy
         


```


b. Read in the fish.png as 'fish'

```{r fish_image}

fish <- "fish.png"


```

c. Make a graph...with gganimate!
```{r gganimate_fish}

aq_plot <- ggplot(aq_tidy, aes(x = as.numeric(year), y=aq_mil, group = country)) + 
  geom_line(aes(color = country)) + 
  geom_point(aes(color = country)) + 
  geom_image(aes(image = fish)) + 
  geom_text(aes(label = country, color = country), position = position_nudge(y = 0.04, x =1), size = 5) + 
  transition_reveal(country, as.numeric(year))

aq_plot

```

##END LAB
